Visual Diagnostics of a\kge{n} \sout{Model} Explainer \kge{Model} -- Tools for the Assessment of LIME Explanations

### 2020-12-05 paper.Rnw
starting to tighten the paper

% Commands
\definecolor{orange}{rgb}{0.94, 0.59, 0.19}
\definecolor{purple}{rgb}{0.5, 0.0, 0.5}
\definecolor{teal}{rgb}{0, 0.502, 0.502}
\newcommand{\hh}[1]{\textcolor{orange}{#1}}
\newcommand{\kgc}[1]{\textcolor{purple}{#1}}
\newcommand{\kge}[1]{\textcolor{teal}{#1}}
\newcommand{\data}{sine data}

@@

% Paper header (title and authors)
\title{Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations}
\author[1]{Katherine Goode*}
\author[1,2]{Heike Hofmann}
\authormark{Goode and Hofmann}
\address[1]{\orgdiv{Department of Statistics}, \orgname{Iowa State University}, \orgaddress{\state{Iowa}, \country{United States}}}
\address[2]{\orgdiv{Center for Statistics and Applications in Forensic Evidence (CSAFE)}, \orgname{Iowa State University}, \orgaddress{\state{Iowa}, \country{United States}}}
\corres{*Katherine Goode, Department of Statistics, Iowa State University, Ames, IA. \email{kgoode@iastate.edu}}

% Abstract and keywords
\abstract[Abstract]{The importance of providing explanations for predictions made by black-box models has led to the development of model explainer methods such as LIME (local interpretable model-agnostic explanations). LIME uses a surrogate model to explain the relationship between predictor variables and predictions from a black-box model in a local region around a prediction of interest. However, the quality of the resulting explanations relies on how well the explainer model captures the black-box model in a specified local region. Here we introduce three visual diagnostics to assess the quality of LIME explanations: (1) explanation scatterplots, (2) assessment metric plots, and (3) feature heatmaps. We apply the visual diagnostics to a forensics bullet matching dataset to show examples where LIME explanations depend on the tuning parameter values and the explainer model oversimplifies the black-box model. Our examples raise concerns about claims made of LIME that are similar to other criticisms in the literature.}
\keywords{explainable machine learning, black-box models, interpretability, statistical graphics, data science}

% Citation information
\jnlcitation{
\cname{\author{Goode K.}, \author{H. Hofmann}, }
\cyear{2020},
\ctitle{Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations},
\cjournal{Stat Anal Data Min: The ASA Data Sci Journal},
\cvol{volume, number and page}.}

@@

In the field of statistics, there are two main uses for models: inference and prediction. Machine learning models are often used for the latter purpose. These models have proven to perform well in a wide range of prediction problems, but the accuracy of many machine learning models comes at the cost of interpretability due to their algorithmic complexity (hence the phrase "black-box models"). Model interpretability allows for the understanding and assessment of how a model produces predictions. The lack of the ability to understand and assess a model makes it difficult to trust the model, especially in areas with high stakes decisions such as medical and forensics sciences. The increased use of machine learning models in applications and the introduction of the General Data Protection Regulation (GDPR) in 2018 \citep{goodman:2016} has resulted in a dramatic increase in explainable machine learning research, which focuses on developing ways to explain output from machine learning algorithms.

@@

Numerous methods have been proposed to provide explanations for black-box model predictions \citep{gilpin:2018, guidotti:2018, ming:2017, molnar:2019}. Some are specific to one type of model (e.g. \citep{simonyan:2013} and \citep{urbanek:2008}), and others are model-agnostic (e.g. \citep{fisher:2018} and \citep{strumbelj:2014}). In this paper, we focus on one specific model-agnostic method: LIME \citep{ribeiro:2016}.

LIME (local interpretable model-agnostic explanations) is a method that uses a surrogate model to relate predictor variables to black-box model predictions (i.e. a model explainer) \citep{ribeiro:2016}. We distinguish between the terms of model explainer and explainer model: by \emph{model explainer} we denote the method for explaining a complex model using a surrogate model, while the \emph{explainer model}, or simply the \emph{explainer}, is the surrogate model.

While some model explainers focus on understanding a model at the global level, LIME claims to provide explanations for individual predictions (local). Additionally, LIME is designed to work with any model (model-agnostic) and to produce easily understandable results (explanations) \citep{ribeiro:2016}. Conceptually, LIME fits a simple (interpretable) model, the explainer model,  meant to capture the behavior of the (complex) black-box model in a local region around a prediction of interest. The  simple model then provides interpretable estimates for variables that most influenced the prediction made by the complex model.

@@

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(
    data = lime_data_gathered,
    mapping = aes(
      x = feature_stnd_value,
      y = prediction,
      size = distance_value,
      alpha = distance_value,
      color = distance
    )
  ) +
  geom_abline(
    data = coefs_data %>% filter(case == "good"),
    mapping = aes(intercept = int, slope = slope),
    size = 1
  ) +
  scale_alpha(range = c(0.3, 1)) +
  scale_color_manual(values = c(NA, "grey30"), guide = "none") +
  geom_point(
    data = prediction_of_interest_gathered,
    mapping = aes(x = feature_value,
                  y = prediction,
                  fill = color),
    color = "black",
    size = 5,
    shape = 23,
    alpha = 0.75
  ) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(
      feature_stnd_value = 1.2,
      prediction = 1.6,
      feature = c("Feature 1", "Feature 2"),
      slope = c(
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "good",
                   feature == "Feature 1") %>%
            pull(slope) %>%
            round(3)
        ),
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "good",
                   feature == "Feature 2") %>%
            pull(slope) %>%
            round(3)
        )
      )
    ),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = ff,
    size = f1_fs / 2.85
  ) +
  labs(
    x = "",
    y = "Black-Box Prediction",
    title = "Conceptual Depiction of a Faithful Local Explainer Model",
    subtitle = paste("Gower Distance Metric Exponent:", good_gower_power),
    fill = "",
    alpha = "Weight",
    size = "Weight"
  ) +
  theme_linedraw(base_family = ff, base_size = f1_fs) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_text(size = f1_fs, color = "black"),
    plot.title = element_text(size = f1_fs)
  ) +
  guides(
    fill = guide_legend(order = 1),
    size = guide_legend(order = 2),
    alpha = guide_legend(order = 2)
  )

@
\caption{A conceptual depiction of a faithful LIME explainer model in the immediate neighborhood of a prediction of interest. The predictions from a hypothetical black-box model are plotted against the standardized values of the two hypothetical predictor variables. The diamond shaped points represent the location of a prediction of interest. The size and opacity of the circular points indicate the weight assigned based on the distance to the prediction of interest computed using the inverse of the Gower distance metric raised to the power of \Sexpr{good_gower_power}. The black lines represent  a weighted ridge regression model used as an explainer model that reasonably captures the relationship between the black-box predictions and the features in a local region around the prediction of interest. That is, the explainer is faithful to the complex model and produces a reasonable explanation that Feature 1 plays a more important role in the prediction of interest than Feature 2 since the magnitude of the slope associated with Feature 1 is larger.}
\label{fig:figure-01}

<<figure-02, out.width = '6.5in', fig.width = 7.95, fig.height = 3.75, warning = FALSE>>=

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(
    data = lime_data_gathered,
    mapping = aes(
      x = feature_stnd_value,
      y = prediction,
      size = distance_value,
      alpha = distance_value,
      color = distance
    )
  ) +
  geom_abline(
    data = coefs_data %>% filter(case == "bad"),
    mapping = aes(intercept = int, slope = slope),
    size = 1
  ) +
  scale_alpha(range = c(0.3, 1)) +
  scale_color_manual(values = c("grey30", NA), guide = "none") +
  geom_point(
    data = prediction_of_interest_gathered,
    mapping = aes(x = feature_value,
                  y = prediction,
                  fill = color),
    color = "black",
    size = 5,
    shape = 23,
    alpha = 0.75
  ) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(
      feature_stnd_value = 1.2,
      prediction = 1.6,
      feature = c("Feature 1", "Feature 2"),
      slope = c(
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "bad",
                   feature == "Feature 1") %>%
            pull(slope) %>%
            round(3)
        ),
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "bad",
                   feature == "Feature 2") %>%
            pull(slope) %>%
            round(3)
        )
      )
    ),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = ff,
    size = f2_fs / 3
  ) +
  labs(
    x = "",
    y = "Black-Box Prediction",
    title = "Conceptual Depiction of an Unfaithful Local Explainer Model",
    subtitle = paste("Gower Distance Metric Exponent:", bad_gower_power),
    fill = "",
    alpha = "Weight",
    size = "Weight"
  ) +
  theme_linedraw(base_family = ff, base_size = f2_fs) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_text(size = f2_fs, color = "black"),
    plot.title = element_text(size = f2_fs)
  ) +
  guides(
    fill = guide_legend(order = 1),
    size = guide_legend(order = 2),
    alpha = guide_legend(order = 2)
  )

@
\caption{A conceptual depiction of an unfaithful local explainer model in the immediate neighborhood of a prediction of interest. A ridge regression model is fit to the same hypothetical black-box model predictions and predictor variables as in \autoref{fig:figure-01}, but the weights are computed using a Gower exponent of 1. The explainer model is unfaithful to the complex model in any immediate neighborhood of the prediction of interest.}
\label{fig:figure-02}
\end{figure*}

\autoref{fig:figure-01} provides a visualization of this conceptual understanding of LIME. The two plots show the predictions from a hypothetical black-box model plotted against the two hypothetical model predictor variables. The diamond shaped points represent the location of the prediction of interest. A Gower distance metric \citep{gower:1971} is used to define locality such that the size of the points represent the inverse of the Gower distance raised to some value and indicate the proximity to the prediction of interest. In the example of \autoref{fig:figure-01}, an exponent of \Sexpr{good_gower_power} is used to emphasize a very local region around the prediction of interest. A ridge regression model weighted by the proximity values is used as the explainer model with the black-box predictions as the response variable and standardized versions of two features in the data as predictor variables. Standardized features allow direct comparisons of the model coefficients. The explainer model is depicted by the black lines in the figure.

The plot on the left of \autoref{fig:figure-01} shows the relationship between the black-box predictions and Feature 1. Here, the explainer model is plotted with Feature 2 fixed at the observed value of the prediction of interest. The explainer model captures the relationship in an  immediate neighborhood around the prediction of interest with a slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 1") %>% pull(slope) %>% round(3)}. The plot on the right shows no global or local relationships between the black-box predictions and Feature 2. Here, the explainer model is plotted with Feature 1 set as the observed value of the prediction of interest, and it has an appropriately small slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 2") %>% pull(slope) %>% round(3)}. The magnitude of the slope associated with Feature 1 is larger than the slope of Feature 2, which suggests that Feature 1 plays a more important role in the prediction made by the black-box model for the prediction of interest. This explanation agrees with a visual assessment of the relationships between the predictions and predictor variables.

The concept of LIME is relatively simple: use an interpretable model to approximate a complex model in a local region. However, a practical implementation of LIME is not straightforward, and research is being done to improve the procedure \citep{laugel:2018}. The current implementations of LIME \citep{ribeiro:2020} \citep{pedersen:2020} offer various tuning parameters (see Section \ref{background}) that affect the explainer model and ultimately, the explanation. Since the explainer model is an approximation of the complex model and not a direct interpretation, the explanations produced by an explainer model are subject to the quality of the approximation. Thus, in order to achieve accurate explanations, the tuning parameter values selected need to be assessed.

We consider an example where the choice of exponent for the weights is crucial to the quality of the explanation. The plots in Figure \ref{fig:figure-02} show the same data as \autoref{fig:figure-01}, but the Gower distance metric exponent is decreased to \Sexpr{bad_gower_power} (the default exponent in the \emph{lime} R package). This causes the observations that are further away from the prediction of interest to be given larger weights than before. In \autoref{fig:figure-01}, the explainer model captures the relationship between the black-box predictions in an immediate neighborhood of the prediction of interest. This cannot be said of the explainer model in \autoref{fig:figure-02}. In addition, the magnitude of the slope associated with Feature 2  (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 2") %>% pull(slope) %>% round(3) %>% abs()}) is larger than that of Feature 1 (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 1") %>% pull(slope) %>% round(3) %>% abs()}). Thus, this explainer model actually provides a misleading explanation that Feature 2 plays a more important role in the prediction of interest.

Several sources in the literature discuss the performance of LIME. One of the biggest difficulties with LIME is determining how to specify a local region \citep{laugel:2018} \citep{molnar:2019}. This is due to an unclear definition of a "local region" and how to apply LIME to achieve an appropriate local region as demonstrated by \autoref{fig:figure-02}. \citet{alvarezmelis:2018} raise a concern pertaining to the robustness of explanations from LIME and other model explainers: they find that even small changes in predictor variables can lead to very different LIME explanations. Additionally, \citet{ribeiro:2016} acknowledge that if a linear model is used as the explainer, LIME relies on a linear approximation of the explainer model to the complex model and state "if the underlying model is highly non-linear even in the locality of the prediction, there may not be a faithful explanation".

As a result of the various ways LIME can fail, it is important to assess LIME explanations. We suggest the use of visual diagnostics for assessment. In this paper, we lay out the set of claims about LIME made by \citet{ribeiro:2016} and propose three visualizations for the assessment of these claims: (1) \emph{explanation scatterplot}, (2) \emph{feature heatmap}, (3) \emph{assessment metric plot}. While LIME is implemented for image, tabular, and text data, we only focus on tabular data. For additional simplicity, we only discuss classification prediction models with a dichotomous response variable and continuous predictor variables. However, the proposed diagnostics may be extended to a wider range of situations.

>>>>

% Commands
\definecolor{orange}{rgb}{0.94, 0.59, 0.19}
\definecolor{purple}{rgb}{0.5, 0.0, 0.5}
\definecolor{teal}{rgb}{0, 0.502, 0.502}
\newcommand{\hh}[1]{\textcolor{orange}{#1}}
\newcommand{\kgc}[1]{\textcolor{purple}{#1}}
\newcommand{\kge}[1]{\textcolor{teal}{#1}}
\newcommand{\data}{sine data}
\renewcommand{\sout}[1]{\unskip}

@@

% Paper header (title and authors)
\title{Visual Diagnostics of a\kge{n} \sout{Model} Explainer \kge{Model} -- Tools for the Assessment of LIME Explanations}
\author[1]{Katherine Goode*}
\author[1,2]{Heike Hofmann}
\authormark{Goode and Hofmann}
\address[1]{\orgdiv{Department of Statistics}, \orgname{Iowa State University}, \orgaddress{\state{Iowa}, \country{United States}}}
\address[2]{\orgdiv{Center for Statistics and Applications in Forensic Evidence (CSAFE)}, \orgname{Iowa State University}, \orgaddress{\state{Iowa}, \country{United States}}}
\corres{*Katherine Goode, Department of Statistics, Iowa State University, Ames, IA. \email{kgoode@iastate.edu}}

% Abstract and keywords
\abstract[Abstract]{The importance of providing explanations for predictions made by black-box models has led to the development of \kge{explainer} model \sout{explainer} methods such as LIME (local interpretable model-agnostic explanations). LIME uses a surrogate model to explain the relationship between predictor variables and predictions from a black-box model in a local region around a prediction of interest. However, the quality of the resulting explanations relies on how well the explainer model captures the black-box model in a specified local region. Here we introduce three visual diagnostics to assess the quality of LIME explanations: (1) explanation scatterplots, (2) assessment metric plots, and (3) feature heatmaps. We apply the visual diagnostics to a forensics bullet matching dataset to show examples where LIME explanations depend on the tuning parameter values and the explainer model oversimplifies the black-box model. Our examples raise concerns about claims made of LIME that are similar to other criticisms in the literature.}
\keywords{explainable machine learning, black-box models, interpretability, statistical graphics, data science}

% Citation information
\jnlcitation{
\cname{\author{Goode K.}, \author{H. Hofmann}, }
\cyear{2020},
\ctitle{Visual Diagnostics of a\kge{n} \sout{Model} Explainer \kge{Model} -- Tools for the Assessment of LIME Explanations},
\cjournal{Stat Anal Data Min: The ASA Data Sci Journal},
\cvol{volume, number and page}.}

@@

In the field of statistics, there are two main uses for models: inference and prediction. Machine learning models \sout{are often used for the latter purpose. These models} have proven to perform well \kge{for problems with the latter objective} \sout{in a wide range of prediction problems}, but the accuracy of many machine learning models comes at the cost of interpretability due to their algorithmic complexity (hence the phrase "black-box models"). Model interpretability allows for the understanding and assessment of how a model produces predictions. The lack of the ability to understand and assess a model makes it difficult to trust the model, especially in areas with high stakes decisions such as \kge{the} medical and forensics sciences. The increased use of machine learning models in applications and the introduction of the General Data Protection Regulation (GDPR) in 2018 \citep{goodman:2016} has resulted in a dramatic increase in explainable machine learning research, which focuses on developing ways to explain output from machine learning algorithms.

@@

Numerous methods have been proposed to provide explanations for black-box model predictions \citep{gilpin:2018, guidotti:2018, ming:2017, molnar:2019}. Some are specific to one type of model (e.g. \citep{simonyan:2013} and \citep{urbanek:2008}), and others are model-agnostic (e.g. \citep{fisher:2018} and \citep{strumbelj:2014}). In this paper, we focus on \kge{the} \sout{one specific} model-agnostic method\sout{:} \kge{of} LIME \citep{ribeiro:2016}.

LIME (local interpretable model-agnostic explanations) \sout{is a method that} uses a surrogate model, to relate predictor variables to black-box model predictions \sout{(i.e. a model explainer)} \citep{ribeiro:2016}. \sout{We distinguish between the terms of model explainer and explainer model: by \emph{model explainer} we denote the method for explaining a complex model using a surrogate model, while the \emph{explainer model}, or simply the \emph{explainer}, is the surrogate model.} While some \kge{explainer} model\kge{s} \sout{explainers} focus on understanding a model at the global level, LIME claims to provide explanations for individual predictions (local). Additionally, LIME is designed to work with any model (model-agnostic) and to produce easily understandable results (explanations) \citep{ribeiro:2016}. Conceptually, LIME fits a simple (interpretable) model, the explainer model, \kge{to approximate} \sout{meant to capture the behavior of} the \sout{(}complex\sout{) black-box} model in a local region around a prediction of interest. The simple model \kge{ma be interpreted to identify the} \sout{then provides interpretable estimates for} variables that most influenced the \kge{complex model} prediction\sout{made by the complex model}.

\kge{While t}\sout{T}he concept of LIME is relatively simple\sout{: use an interpretable model to approximate a complex model in a local region. However}, a practical implementation of LIME is not straightforward, and research is being done to improve the procedure \citep{laugel:2018}. The current implementations of LIME \citep{ribeiro:2020} \citep{pedersen:2020} offer various tuning parameters (see Section \ref{background}) that affect the explainer model and ultimately, the explanation. Since the explainer model is an approximation of the complex model and not a direct interpretation, the explanations produced by an explainer model are subject to the quality of the approximation. Thus, in order to achieve accurate explanations, the tuning parameter values selected need to be assessed.

\kge{To demonstrate the effect of tuning parameters on LIME, we consider two different explainer models applied to the same prediction. Suppose a black-box model is trained using two features: Feature 1 and Feature 2. Figure \ref{fig:figure-01} shows the black-box model predictions versus standardized versions of Feature 1 (left) and Feature 2 (right). The diamond shaped point in the plots indicates the location of the prediction of interest to explain. A weight ridge regression with the black-box predictions as the response variable and standardized versions the features as predictor variables will be used as the explainer model in both scenarios, but the weights will vary. Note that the standardized features allow for direct comparisons of the model coefficients.}

@@

\kge{Figure \ref{fig:figure-01} depicts the first scenario where the weights are computed as the inverse proximity of the point to the prediction of interest computed using the Gower distance metric \citep{gower:1971} raised to the power of \Sexpr{good_gower_power} to emphasize a very local region. The point size represents the weight, and the black lines depict the ridge regression model given the observed prediction of interest value from the opposite feature. For Feature 1 (left), the explainer model captures the relationship in an immediate neighborhood around the prediction of interest with a slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 1") %>% pull(slope) %>% round(3)}, and for Feature 2 (right), the ridge regression has a small slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 2") %>% pull(slope) %>% round(3)} that accurately reflects the absence of any global or local relationship between the black-box predictions and Feature 2. The magnitude of the slope associated with Feature 1 is larger than the slope of Feature 2, which suggests that Feature 1 plays a more important role in the prediction made by the black-box model for the prediction of interest. This explanation agrees with a visual assessment of the relationships between the predictions and predictor variables.}

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(
    data = lime_data_gathered,
    mapping = aes(
      x = feature_stnd_value,
      y = prediction,
      size = distance_value,
      color = distance
    ), alpha = 0.6
  ) +
  geom_abline(
    data = coefs_data %>% filter(case == "good"),
    mapping = aes(intercept = int, slope = slope),
    size = 1
  ) +
  scale_color_manual(values = c(NA, "grey30"), guide = "none") +
  geom_point(
    data = prediction_of_interest_gathered,
    mapping = aes(x = feature_value,
                  y = prediction,
                  fill = color),
    color = "black",
    size = 5,
    shape = 23,
    alpha = 0.75
  ) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(
      feature_stnd_value = 1.2,
      prediction = 1.6,
      feature = c("Feature 1", "Feature 2"),
      slope = c(
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "good",
                   feature == "Feature 1") %>%
            pull(slope) %>%
            round(3)
        ),
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "good",
                   feature == "Feature 2") %>%
            pull(slope) %>%
            round(3)
        )
      )
    ),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = ff,
    size = f1_fs / 2.85
  ) +
  labs(
    x = "",
    y = "Black-Box Prediction",
    title = "Conceptual Depiction of a Faithful Local Explainer Model",
    subtitle = paste("Gower Distance Metric Exponent:", good_gower_power),
    fill = "",
    alpha = "Weight",
    size = "Weight"
  ) +
  theme_linedraw(base_family = ff, base_size = f1_fs) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_text(size = f1_fs, color = "black"),
    plot.title = element_text(size = f1_fs)
  ) +
  guides(
    fill = guide_legend(order = 1),
    size = guide_legend(order = 2),
    alpha = guide_legend(order = 2)
  )

@
\caption{A conceptual depiction of a faithful LIME explainer model in the immediate neighborhood of a prediction of interest. \kge{The explainer model in this scenario is a ridge regression weighted by the inverse Gower distance to the prediction of interest raised to the power of \Sexpr{good_gower_power}.} \sout{The predictions from a hypothetical black-box model are plotted against the standardized values of the two hypothetical predictor variables. The diamond shaped points represent the location of a prediction of interest. The size and opacity of the circular points indicate the weight assigned based on the distance to the prediction of interest computed using the inverse of the Gower distance metric raised to the power of \Sexpr{good_gower_power}. The black lines represent a weighted ridge regression model used as an explainer model that reasonably captures the relationship between the black-box predictions and the features in a local region around the prediction of interest. That is, the explainer is faithful to the complex model and produces a reasonable explanation that Feature 1 plays a more important role in the prediction of interest than Feature 2 since the magnitude of the slope associated with Feature 1 is larger.}}
\label{fig:figure-01}
\end{figure*}

\sout{\autoref{fig:figure-01} provides a visualization of this conceptual understanding of LIME. The two plots show the predictions from a hypothetical black-box model plotted against the two hypothetical model predictor variables. The diamond shaped points represent the location of the prediction of interest. A Gower distance metric \citep{gower:1971} is used to define locality such that the size of the points represent the inverse of the Gower distance raised to some value and indicate the proximity to the prediction of interest. In the example of \autoref{fig:figure-01}, an exponent of \Sexpr{good_gower_power} is used to emphasize a very local region around the prediction of interest. A ridge regression model weighted by the proximity values is used as the explainer model with the black-box predictions as the response variable and standardized versions of two features in the data as predictor variables. Standardized features allow direct comparisons of the model coefficients. The explainer model is depicted by the black lines in the figure.}

\sout{The plot on the left of \autoref{fig:figure-01} shows the relationship between the black-box predictions and Feature 1. Here, the explainer model is plotted with Feature 2 fixed at the observed value of the prediction of interest. The explainer model captures the relationship in an immediate neighborhood around the prediction of interest with a slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 1") %>% pull(slope) %>% round(3)}. The plot on the right shows no global or local relationships between the black-box predictions and Feature 2. Here, the explainer model is plotted with Feature 1 set as the observed value of the prediction of interest, and it has an appropriately small slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 2") %>% pull(slope) %>% round(3)}. The magnitude of the slope associated with Feature 1 is larger than the slope of Feature 2, which suggests that Feature 1 plays a more important role in the prediction made by the black-box model for the prediction of interest. This explanation agrees with a visual assessment of the relationships between the predictions and predictor variables.}

\sout{The concept of LIME is relatively simple: use an interpretable model to approximate a complex model in a local region. However, a practical implementation of LIME is not straightforward, and research is being done to improve the procedure \citep{laugel:2018}. The current implementations of LIME \citep{ribeiro:2020} \citep{pedersen:2020} offer various tuning parameters (see Section \ref{background}) that affect the explainer model and ultimately, the explanation. Since the explainer model is an approximation of the complex model and not a direct interpretation, the explanations produced by an explainer model are subject to the quality of the approximation. Thus, in order to achieve accurate explanations, the tuning parameter values selected need to be assessed.}

\kge{Figure \ref{fig:figure-02} depicts the second explainer model scenario. Again, the weights are computed using the Gower distance metric, but in this situation, the distances are raise to a power of \Sexpr{bad_gower_power} (the default exponent in the \emph{lime} R package).} \sout{We consider an example where the choice of exponent for the weights is crucial to the quality of the explanation. The plots in Figure \ref{fig:figure-02} show the same data as \autoref{fig:figure-01}, but the Gower distance metric exponent is decreased to \Sexpr{bad_gower_power} (the default exponent in the \emph{lime} R package).} This causes the observations that are further away from the prediction of interest to be given larger weights than before. In \autoref{fig:figure-01}, the explainer model captures the relationship between the black-box predictions in an immediate neighborhood of the prediction of interest. This cannot be said of the explainer model in \autoref{fig:figure-02}. In addition, the magnitude of the slope associated with Feature 2  (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 2") %>% pull(slope) %>% round(3) %>% abs()}) is larger than that of Feature 1 (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 1") %>% pull(slope) %>% round(3) %>% abs()}). Thus, this explainer model actually provides a misleading explanation that Feature 2 plays a more important role in the prediction of interest.

\begin{figure*}[!thp]
<<figure-02, out.width = '6.5in', fig.width = 7.95, fig.height = 3.75, warning = FALSE>>=

@@

# Plot of good explainer model
ggplot() +
  facet_grid(. ~ feature, switch = "x") +
  geom_point(
    data = lime_data_gathered,
    mapping = aes(
      x = feature_stnd_value,
      y = prediction,
      size = distance_value,
      color = distance
    ), 
    alpha = 0.6
  ) +
  geom_abline(
    data = coefs_data %>% filter(case == "bad"),
    mapping = aes(intercept = int, slope = slope),
    size = 1
  ) +
  scale_color_manual(values = c("grey30", NA), guide = "none") +
  geom_point(
    data = prediction_of_interest_gathered,
    mapping = aes(x = feature_value,
                  y = prediction,
                  fill = color),
    color = "black",
    size = 5,
    shape = 23,
    alpha = 0.75
  ) +
  scale_fill_manual(values = "#FAAA72") +
  geom_text(
    data = data.frame(
      feature_stnd_value = 1.2,
      prediction = 1.6,
      feature = c("Feature 1", "Feature 2"),
      slope = c(
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "bad",
                   feature == "Feature 1") %>%
            pull(slope) %>%
            round(3)
        ),
        paste(
          "Slope:",
          coefs_data %>%
            filter(case == "bad",
                   feature == "Feature 2") %>%
            pull(slope) %>%
            round(3)
        )
      )
    ),
    mapping = aes(x = feature_stnd_value,
                  y = prediction,
                  label = slope),
    family = ff,
    size = f2_fs / 3
  ) +
  labs(
    x = "",
    y = "Black-Box Prediction",
    title = "Conceptual Depiction of an Unfaithful Local Explainer Model",
    subtitle = paste("Gower Distance Metric Exponent:", bad_gower_power),
    fill = "",
    alpha = "Weight",
    size = "Weight"
  ) +
  theme_linedraw(base_family = ff, base_size = f2_fs) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.placement = "outside",
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.x = element_text(size = f2_fs, color = "black"),
    plot.title = element_text(size = f2_fs)
  ) +
  guides(
    fill = guide_legend(order = 1),
    size = guide_legend(order = 2),
    alpha = guide_legend(order = 2)
  )

@
\caption{A conceptual depiction of an unfaithful local explainer model in the immediate neighborhood of a prediction of interest. A\kge{gain, a weighted} ridge regression \kge{is used as the explainer model} \sout{is fit to the same hypothetical black-box model predictions and predictor variables as in \autoref{fig:figure-01}}, but the \kge{Gower distance is raised to a power} \sout{weights are computed using a Gower exponent} of 1. \sout{The explainer model is unfaithful to the complex model in any immediate neighborhood of the prediction of interest.}}
\label{fig:figure-02}
\end{figure*}

\kge{This example demonstrates the effect of the selection of weights on the LIME explanations, but other concerns with LIME have been raised in the literature.} \sout{Several sources in the literature discuss the performance of LIME.} One of the biggest difficulties with LIME is determining how to specify a local region \citep{laugel:2018} \citep{molnar:2019}. This is due to an unclear definition of a "local region" and how to apply LIME to achieve an appropriate local region as demonstrated by \autoref{fig:figure-02}. \citet{alvarezmelis:2018} raise a concern pertaining to the robustness of explanations from LIME and other \kge{explainer} model\kge{s} \sout{explainers}: they find that even small changes in predictor variables can lead to very different LIME explanations. Additionally, \citet{ribeiro:2016} acknowledge that if a linear model is used as the explainer, LIME relies on a linear approximation of the explainer model to the complex model and state "if the underlying model is highly non-linear even in the locality of the prediction, there may not be a faithful explanation".

\kge{In this paper, we stress the importance of assessing LIME by introducing several diagnostic plots that highlight key issues with LIME.} \sout{As a result of the various ways LIME can fail, it is important to assess LIME explanations. We suggest the use of visual diagnostics for assessment. In this paper,} \kge{To do this,} we lay out the set of claims about LIME made by \citet{ribeiro:2016} and propose three visualizations for the assessment of these claims: (1) \emph{explanation scatterplots}, (2) \emph{feature heatmaps}, (3) \emph{assessment metric plots}. While LIME is implemented for image, tabular, and text data, we only focus on tabular data. For additional simplicity, we only discuss classification prediction models with a dichotomous response variable and continuous predictor variables. However, the proposed diagnostics may be extended to \kge{other} \sout{a wider range of} situations.
